---
title: "Datasaurus Analysis"
author: "Prabuddha Deore"
date: "2025-05-09"
output: pdf_document
---

```{r}
library(tidyverse) 
library(datasauRus)

```
```{r}
#?datasaurus_dozen

```
```{r}
# Count the rows and columns and list the variables
dim(datasaurus_dozen)
names(datasaurus_dozen)

```

```{r}
# Filter for the 'dino' dataset
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

# Preview the first few rows
head(dino_data)
```
```{r}
# Compute summary statistics
summary_stats <- dino_data %>%
  summarise(
    mean_x = mean(x),
    mean_y = mean(y),
    sd_x = sd(x),
    sd_y = sd(y),
    correlation = cor(x, y)
  )

# Print the summary
summary_stats

```
```{r}
ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

```{r all-datasets-plot, fig.width=10, fig.height=15}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset))+
  geom_point(size= 1.5)+
  facet_wrap(~ dataset, ncol = 3) +
  theme(legend.position = "none")
```

## Summary

The Datasaurus Dozen data set highlights the importance of data visualization. While the summary statistics, such as mean, standard deviation, and correlation, are very similar for these data sets, their visual forms show very different shapes and patterns. This fact demonstrates that using only statistical measures can lead to incorrect interpretations and highlights the importance of using visualizations to identify underlying structures or anomalies in the data.

